\section{Discussion}
\label{sec:discussion}

\para{What do the results mean?} The main empirical result is a qualified answer to the motivating question. Interpolation between incongruent concepts does not have one generic signature. Instead, instability depends on the kind of mismatch that defines the endpoint pair. When the mismatch changes relational structure, the path behaves more like the geodesic failures described in latent-space geometry work \cite{shao2017riemannian,michelis2021linear}. When the mismatch is a short lexical substitution, the negative endpoint can remain close enough to the source that the linear path still looks locally regular.

\para{Why does pooled incongruence look easier?} The pooled reversal comes from benchmark construction, not from evidence against path sensitivity. In two of the three subsets, the negative caption is often lexically closer to the source than the paraphrase. That makes the incongruent endpoint a shorter chord in embedding space. A shorter chord naturally reduces nearest-neighbor deviation and dual-space travel, even if the endpoint is less truthful. The result is a reminder that truth-conditional corruption and geometric difficulty are not the same variable.

\para{Broader implications.} These findings narrow when linear traversal is a useful analysis tool. If a representation change preserves local lexical scaffolding, a straight path may remain interpretable even when the endpoint is wrong. But if the change disrupts who did what to whom, path-based methods become less reliable. This matters for activation steering, latent editing, and causal analysis, where a direction can look clean at the endpoints while the intermediate states are semantically unstable.

\para{Limitations.} Our study uses embedding-space geometry rather than hidden-state geometry with the causal inner product of Park \etal \cite{park2024lrh}. The results therefore test a broader LRH-inspired hypothesis rather than the full formalism. The graph detour metric is also weak in this setting because the sampled $k$-NN graph produces a detour ratio of 1.0 for every example, leaving off-manifold and dual-space measures as the informative diagnostics. Finally, \sugarcrepe is only a proxy for incongruence. Its subsets mix distinct corruption types, and the original \texttt{swap\_atribute} split is orthographically inconsistent and semantically closer to lexical editing than to structural contradiction.
